---
title: "Age structure for commune level"
author: "manh"
date: '2023-03-20'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Introduction

The code below uses raster-based modelled data from WorldPop on age structure to
generate a data frame of age structure by commune level

## Local folders

```{r folders}
gadm_data_folder <- "~/OneDrive - Oxford University Clinical Research Unit/DART/GADM/"
wpop_data_folder <- "/Volumes/special$/DART/GADM\ age\ structure" ## I assumed you have data somewhere else
```

## Packages

Packages needed for the analysis:

```{r}
needed_package <- c("dplyr", "tidyr", "purrr", "terra", "stringr", "sf", "parallel", "magrittr")
```

Installing the packages that are not already installed:

```{r}
to_install <- needed_package[which(!needed_package %in% installed.packages()[, "Package"])]
if (length(to_install)) install.packages(to_install)
```

Loading the packages:

```{r message = FALSE}
dev_null <- sapply(needed_package, function(libname) library(libname, character.only = TRUE))
rm(dev_null)
```


## Downloading commune polygons data from GADM

GADM root URL:

```{r}
gadm <- "https://geodata.ucdavis.edu/gadm/"
```

Version and format:

```{r}
vers_form <- "gadm3.6/Rsf/"
```

File:

```{r}
file_name <- "gadm36_VNM_3_sf.rds"
```

Creating the local folder structure if it does not exist:

```{r}
local_folder <- paste0(gadm_data_folder, vers_form)
if (!dir.exists(local_folder)) dir.create(local_folder, recursive = TRUE)
```

Downloading the data if not already available locally:

```{r}
local_file <- paste0(local_folder, file_name)
if (!file.exists(local_file)) download.file(paste0(gadm, vers_form, file_name), local_file)
```

Loading the polygons data.

```{r}
communes <- local_file %>%
  readRDS() %>%
  mutate(
    NAME_1 = stringi::stri_trans_general(NAME_1, "any-ascii"),
    NAME_2 = stringi::stri_trans_general(NAME_2, "any-ascii")
  ) %>%
  mutate(commune = paste0(NAME_1, ",", NAME_2, ",", VARNAME_3)) %>% ## a commune is identified with province, district and commune
  st_set_crs(4326)
```

Checking the map of the provinces polygons:

```{r}
communes %>%
  st_geometry() %>%
  plot()
```

All the combinations of years, ages and gender:

```{r}
ages <- c(0, 1, seq(5, 80, 5))

combinations <- expand_grid(
  years = 2000:2020,
  ages = ages,
  sexes = c("f", "m")
)
```

generate age structure commune level 
```{r}
commune2vect <- function(commune_name) {
  communes %>%
    filter(commune == commune_name) %>%
    vect()
}
```
A function that crops and masks a `SpatRaster` object `rst` with a `SpatVector`
object `pol`:

```{r}
crop_mask <- function(rst, pol) {
  rst %>%
    crop(pol) %>%
    mask(pol)
}
```

A function that extracts the number of people of a raster `rst` inside a polygon
`pol`:

```{r}
nb_people <- function(rst, pol) {
  crop_mask(rst, pol) %>%
    values() %>%
    sum(na.rm = TRUE)
}
```
A function that converts a TIFF file (i.e. number of people for a given year,
age, and sex) into a data frame with number of people per commune:

```{r}
tif2df <- function(tif_file) {
  tmp <- first(str_split(str_remove(tif_file, ".tif"), "_"))
  tif_file %>%
    paste0(wpop_data_folder, "/", .) %>%
    rast() %>%
    map_dbl(commune_vect_list, nb_people, rst = .) %>%
    tibble(year = tmp[4], commune = commune_names, sex = tmp[2], age = tmp[3], n = .)
}
```

The names of the each commune (follow by province and district names)

```{r}
commune_names <- communes$commune
```

A list of provinces polygons as `SpatVector`:

```{r}
commune_vect_list <- map(commune_names, commune2vect)
```

 The world population data is stored on OUCRU shared folder. These functions takes quite a while to run might be of connection issues. I did not put below code in a function and run the code for a numbers of .tif files instead
 
```{r}
list_file <- dir(wpop_data_folder)

age_data <- list_file[1] %>%
  mclapply(tif2df, mc.cores = detectCores() - 1) %>%
  bind_rows() %>%
  mutate(
    age_class = hash[age],
    sex = c(f = "female", m = "male")[sex],
    age = as.integer(age)
  )

saveRDS(age_data, "~/OneDrive - Oxford University Clinical Research Unit/DART/WorldPop/processed/age structure/Vietnam/age_data_commune/age_data_commune_1")
```
